A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
metrics

\newcommand{\eval}{\c{eval}\xspace}
\newcommand{\parse}{\c{parse}\xspace}
\newcommand{\source}{\c{source}\xspace}
\newcommand{\local}{\c{local}\xspace}
\newcommand{\unlockBinding}{\c{unlockBinding}\xspace}
\newcommand{\substitute}{\c{substitute}\xspace}
\newcommand{\datatable}{\c{data.table}\xspace}
\newcommand{\mlogit}{\c{mlogit}\xspace}
\newcommand{\mboost}{\c{mboost}\xspace}
\newcommand{\metafor}{\c{metafor}\xspace}
\newcommand{\lavaan}{\c{lavaan}\xspace}
\newcommand{\mclust}{\c{mclust}\xspace}
\newcommand{\gamlss}{\c{gamlss}\xspace}
\newcommand{\ggproto}{\c{ggproto}\xspace}
\newcommand{\ggplot}{\c{ggplot2}\xspace}

\newcommand{\base}{\c{base}\xspace}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

We have selected the TOP \CorpusPackages packages based on their reverse
dependencies, \ie how many other packages depend on them.\footnote{Extracted
  from package metadata using a builtin function.}  This helps us find
packages which should be better written and have better code coverage.  The
resulting set contains \CorpusRCodeRnd lines of R code and
\CorpusNativeCodeRnd lines of native code.\footnote{Excluding comments and
  blank lines using \url{github.com/AlDanial/cloc}.}  For each package we
use its runnable code to compute the package code coverage\footnote{Computed
  using \url{ https://github.com/r-lib/covr}}. On average, across the
\CorpusPackages packages, it is \CorpusMeanExprCoverage.
Figure~\ref{fig:corpus} shows these packages, the size of the dots reflects
the project's size in lines of code.  The x-axis indicates code coverage in
percents and the y-axis gives the number of call sites to \eval in log
scale. Dotted lines indicate means. Packages with over
\CorpusEvalsPackageTreshold eval call sites are named.

All included packages come from the Comprehensive R Archive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \CorpusAllCranRnd packages\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}}. Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality. Most relevant for this work is that all of
the \emph{runnable code} is tested and only a successfully running package is
admitted in the archive. This is important as we use this runnable code for the
dynamic analysis of eval calls.

@@

To represent end-user code in the corpus, we turned to
Kaggle\footnote{\url{https://www.kaggle.com}}, an online platform for
data-science and machine-learning. The website allow people to share
data-science competitions, a data-analysis problem together with data, for
which users try to find the best solution. The solutions, called {kernels},
are then shared back to the platform in the form of either plain scripts or
as notebooks.  One of the most popular competition is about predicting
passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

@@

This section presents an overview of the usage of \eval in our corpus.
Previous work reported that 82\% of JavaScript-enabled web pages used
\eval~\cite{ecoop11}. The case for R is more striking with 100\% of the
programs using that \eval.\footnote{In this section, we combine calls to
  \c{eval}, \c{eval.parent}, \c{evalq} and \c{local}.}  This high number is
explained by the fact that the R implementation, its core packages, use
\eval extensively. In our analysis, we differentiate uses of \eval based on
where they originate from. A {\bf core} \eval is one that occurs in the code
of a core package.  A {\bf dependent} \eval is one that occurs in a package
that was included by the target package. An {\bf own} \eval is one that
occurs in the source code of the target package.

Over all, we observed \AllAllCallCountRnd calls to \eval in the
\CorpusAllProgramsRnd programs in our corpus.  Table~\ref{A} details this
run-time information. The format is T (m/M/x) where T is the total number of
calls observed for all runs of the corpus, m and x are the minimum and
maximum number of calls for any given run, and M is the mean number of calls
per run. We can observe that an average run of CRAN package will trigger 581
{\eval}s in core package, 532 in dependent packages and 292 in the package
itself.  The Kaggle data set has small programs, and, these programs only
trigger \eval in core and dependent packages. Thus, an average Kaggle run
has 412 core calls, 96 dependent calls, and no own calls.

 \begin{table*}[tb]
   \centering
   \begin{tabular}{r|c|c|c@{}} \hline
     Corpus / Source & Core                              & Dependent                              & Own \\\hline
     CRAN   & \summaryrnd{CranCoreEvalCalls}    & \summaryrnd{CranDependentEvalCalls}    & \summaryrnd{CranOwnEvalCalls} \\
     Kaggle & \summaryrnd{KaggleCoreEvalCalls}  & \summaryrnd{KaggleDependentEvalCalls}  & \summaryrnd{KaggleOwnEvalCalls} \\\hline
   \end{tabular}
   \caption{Eval calls in CRAN and Kaggle}
   \label{A}
 \end{table*}

%% The following table will be replaced. {\bf <At this point separating between
%%   the different kinds of evals is distracting. We have not argued why it is
%%   important.>}

%% \begin{table}[ht]\resizebox{\columnwidth}{!}{\begin{tabular}{l|rr|r||rr|r}\hline
%% \multicolumn{1}{c}{Function}& \multicolumn{3}{c}{Calls} & \multicolumn{3}{c}{Callsites}\\
%% \cline{1-1} \cline{2-4} \cline{5-7}
%% & Core & Packages & Total & Core & Packages & Total\\\hline
%% \c{eval} & \CoreEvalCallCountRnd & \PackageEvalCallCountRnd  & \AllEvalCallCountRnd & \CoreEvalSiteCountRnd & \PackageEvalSiteCountRnd &  \AllEvalSiteCountRnd\\
%% \c{eval.parent} & \CoreEvalParentCallCountRnd & \PackageEvalParentCallCountRnd  & \AllEvalParentCallCountRnd & \CoreEvalParentSiteCountRnd & \PackageEvalParentSiteCountRnd & \AllEvalParentSiteCountRnd\\
%% \c{evalq} & \CoreEvalqCallCountRnd & \PackageEvalqCallCountRnd  & \AllEvalqCallCountRnd & \CoreEvalqSiteCountRnd & \PackageEvalqSiteCountRnd & \AllEvalqSiteCountRnd \\
%% \c{local} & \CoreLocalCallCountRnd & \PackageLocalCallCountRnd  & \AllLocalCallCountRnd & \CoreLocalSiteCountRnd & \PackageLocalSiteCountRnd & \AllLocalSiteCountRnd \\\hline
%% Total & \CoreAllCallCountRnd & \PackageAllCallCountRnd  & \AllAllCallCountRnd & \CoreAllSiteCountRnd & \PackageAllSiteCountRnd & \AllAllSiteCountRnd \\  \hline
%% \end{tabular} } \label{table:eval-count-summary}
%%  \caption{Distribution of eval calls and callsites in the corpus}
%% \end{table}

%% \ref{table:top-ten-package-summary} shows the number of calls to eval
%% made by the ten most frequent callers to eval. These packages together have
%% \missingNumber callsites and account for \missingPercentage calls to eval.
%% Furthermore, the top ten callsites account for \missingPercentage of package
%% eval calls. A single callsite of \TopTenPackageNameA accounts for over
%% \missingPercentage calls to eval.

%% \begin{table}[ht]
%%   \resizebox{\columnwidth}{!}{%
%%     \begin{tabular}{l|rrr||l|rrr}
%%       \hline
%%       \multicolumn{4}{c}{}& \multicolumn{4}{c}{} \\
%%       \cline{1-4} \cline{5-8}
%%       Package & Callsites & Calls & Calls \% & Package & Callsites & Calls & Calls \%\\
%%       \hline
%%       \TopTenPackageNameA & \TopTenPackageCallsiteCountA & \TopTenPackageCallCountA &\TopTenPackageCallPercA & \TopTenPackageNameF & \TopTenPackageCallsiteCountF & \TopTenPackageCallCountF & \TopTenPackageCallPercF \\
%%       \TopTenPackageNameB & \TopTenPackageCallsiteCountB & \TopTenPackageCallCountB &\TopTenPackageCallPercB & \TopTenPackageNameG & \TopTenPackageCallsiteCountG & \TopTenPackageCallCountG & \TopTenPackageCallPercG \\
%%       \TopTenPackageNameC & \TopTenPackageCallsiteCountC & \TopTenPackageCallCountC &\TopTenPackageCallPercC & \TopTenPackageNameH & \TopTenPackageCallsiteCountH & \TopTenPackageCallCountH & \TopTenPackageCallPercH \\
%%       \TopTenPackageNameD & \TopTenPackageCallsiteCountD & \TopTenPackageCallCountD &\TopTenPackageCallPercD & \TopTenPackageNameI & \TopTenPackageCallsiteCountI & \TopTenPackageCallCountI & \TopTenPackageCallPercI \\
%%       \TopTenPackageNameE & \TopTenPackageCallsiteCountE & \TopTenPackageCallCountE &\TopTenPackageCallPercE & \TopTenPackageNameJ & \TopTenPackageCallsiteCountJ & \TopTenPackageCallCountJ & \TopTenPackageCallPercJ \\
%%       \hline
%%     \end{tabular}
%%   }
%%   \label{table:top-ten-package-summary}
%%   \caption{Distribution of eval calls and callsites in the top ten packages}
%% \end{table}

Table \ref{table:site-package-summary} shows the results of static analysis
for the CRAN packages in our corpus. 89 packages have a single call site,
and one package has \miss{252} call sites of \eval. \miss{XXX} packages have
no calls to \eval. The core packages have \miss{41} callsites of eval and
Kaggle has none. These static numbers are under approximations as \eval may
be called reflectively, through aliases, or passed to a higher-order function.

@@

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4 nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. Most large AST originate from library \emph{VGAM}, which is used for statistical modelling. It uses \eval to evaluate large matrices with values in a given environment.

\begin{figure}[htbp]
	\centering
\includegraphics[width=\columnwidth]{ast_sizes}
\caption{Input sizes} \label{fig:ast-size}
\end{figure}

@@

We compare the number of events intercepted during execution of \eval across the
whole corpus. For this, we use the number of interpreter's evaluator loops as a
proxy for events. In the corpus, we observe \AllEventCountRnd events. Only
\EvalEventCountRnd events occur inside eval. This is just \EvalEventAllPerc of
all events. Table~\ref{table:event-distribution} shows the distribution of events
caused by \eval calls from R Core and CRAN packages.

\begin{table}[ht]\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}\hline
Events & Core Eval Call \% & Package Eval Call \% \\ \hline
\EventsMinRangeA--\EventsMaxRangeA & \EventsCoreEvalPercA & \EventsPackageEvalPercA \\
\EventsMinRangeB--\EventsMaxRangeB & \EventsCoreEvalPercB & \EventsPackageEvalPercB \\
\EventsMinRangeC--\EventsMaxRangeC & \EventsCoreEvalPercC & \EventsPackageEvalPercC \\ \hline
\end{tabular}} \label{table:event-distribution}
\caption{Distribution of events in \eval calls across Core and Package}
\end{table}

>>>>

\newcommand{\eval}{\c{eval}}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}

\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

We have selected the TOP \CorpusPackages packages based on their reverse
dependencies, \ie how many other packages depend on them.\footnote{Extracted
  from package metadata using a builtin function.}  This helps us find
packages which should be better written and have better code coverage.  The
resulting set contains \CorpusRCodeRnd lines of R code and
\CorpusNativeCodeRnd lines of native code.\footnote{Excluding comments and
  blank lines using \url{github.com/AlDanial/cloc}.}  For each package we
use its runnable code to compute the package code coverage.\footnote{Computed
  using \url{ https://github.com/r-lib/covr}.} On average, across the
\CorpusPackages packages, it is \CorpusMeanExprCoverage.
Figure~\ref{fig:corpus} shows these packages, the size of the dots reflects
the project's size in lines of code.  The x-axis indicates code coverage in
percents and the y-axis gives the number of call sites to \eval in log
scale. Dotted lines indicate means. Packages with over
\CorpusEvalsPackageTreshold eval call sites are named.

All included packages come from the Comprehensive R Archive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \CorpusAllCranRnd packages.\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}.} Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality. Most relevant for this work is that all of
the \emph{runnable code} is tested and only a successfully running package is
admitted in the archive. This is important as we use this runnable code for the
dynamic analysis of eval calls.

@@

To represent end-user code in the corpus, we turned to
Kaggle,\footnote{\url{https://www.kaggle.com}} an online platform for
data-science and machine-learning. The website allow people to share
data-science competitions, a data-analysis problem together with data, for
which users try to find the best solution. The solutions, called {kernels},
are then shared back to the platform in the form of either plain scripts or
as notebooks.  One of the most popular competition is about predicting
passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

@@

This section presents an overview of the usage of \eval in our corpus.
Richards et al. reported 82\% of JavaScript-enabled web pages using
\eval~\cite{ecoop11}. The case for R is more striking with 100\% of the
programs call the \eval function.\footnote{We combine calls to \c{eval},
  \c{eval.parent}, \c{evalq} and \c{local}.}  This high number is explained
by the fact that the R implementation, its core packages, use \eval
extensively. We differentiate uses based on their origin. A {\bf core} \eval
is one that occurs in the code of a core package.  A {\bf dependent} \eval
occurs in a package imported by the target package. An {\bf own} \eval is
one that occurs in the source code of the target package.

Over all, we observe \AllAllCallCountRnd calls to \eval in our corpus.
Table~\ref{A} details this run-time information. The format is T (m/M/x)
where T is the total number of calls observed for all runs of the corpus, m
and x are the minimum and maximum number of calls for any given run, and M
is the mean number of calls per run. An average CRAN package run triggers
581 {\eval} calls in core package, 532 in dependent packages and 292 in the
CRAN package itself.  The Kaggle data set has small programs which only
trigger \eval in core and dependent packages. Thus, an average Kaggle run
has 412 core calls, 96 dependent calls, and no own calls.

%% SORRY  I HAD TO REFORMAT
\begin{table}[tb] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. 61 packages have a single call site, and one package has 142
call sites of \eval. 60 packages have no calls to \eval. The core packages
%%%% FIX THESE NUMBERS
have \miss{41} callsites of eval and Kaggle has none. These numbers are
under approximations as \eval may be called reflectively, through aliases,
or passed to a higher-order function.

@@

\begin{figure}[!b]
	\centering
\includegraphics[width=\columnwidth]{ast_sizes}
\caption{Input sizes} \label{fig:ast-size}
\end{figure}

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent
of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4
nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. Most large AST originate
from library \c{VGAM}, which is used for statistical modelling. It uses
\eval to evaluate large matrices with values in a given environment.

We compare the number of events intercepted during execution of \eval across
the whole corpus. For this, we use the number of interpreter steps as a
proxy for events. We observe \AllEventCountRnd events all told, and
\EvalEventCountRnd events occur inside eval (\EvalEventAllPerc of all
events). Table~\ref{table:event-distribution} shows the distribution of
events caused by \eval calls.

\begin{table}[ht]\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}\hline
Events & Core Eval Call \% & Package Eval Call \% \\ \hline
\EventsMinRangeA--\EventsMaxRangeA & \EventsCoreEvalPercA & \EventsPackageEvalPercA \\
\EventsMinRangeB--\EventsMaxRangeB & \EventsCoreEvalPercB & \EventsPackageEvalPercB \\
\EventsMinRangeC+ & \EventsCoreEvalPercC & \EventsPackageEvalPercC \\ \hline
\end{tabular}} \label{table:event-distribution}
  \caption{Events in \eval calls}
\end{table}

@@
